Meta-analysis

### 2020-06-19 04_MetaAnalysis.Rmd
I finally understood why we divide by n to get sample variance from mean squared error. And some other tidying up

### Caluculate standardised mean differences
We may wish to undertake the meta-analysis using standardised mean differences. 
Lets calculate them and add them to the dataframe.

```{r remove_studies}
#slimmer_PM_dat <-
  # slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),]
```

@@

```{r MSE_2_SVar, eval=FALSE, include=FALSE}
#calculation when using log of response
slimmer_PM_dat$vi <-
   slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
                                  2)

@@

```{r}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for(T_ref in slimmer_PM_dat$trial_ref){

@@

@@

```{r metafor_organisation}
#remove controls from the data
dat1 <- slimmer_PM_dat[!slimmer_PM_dat$fungicide_ai == "control",]

dat1$spray_management <-
   factor(
      dat1$spray_management,
      c( "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   )

@@

Lets undertake the meta-analysis, we are using the `spray_management` variable as a moderator and a interactive term to the `trial` variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.
```{r Metafor-analysis}
dat1$vi <- dat1$yield_error

PM_mv <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management -1,
   method = "ML",
   random = ~ spray_management|trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
In this result we can see that the `Early` treatment is not significantly different to zero, which in this case is the mean of the no spray control. 
However the other treatments are significantly different to zero.
The Qm [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, show the moderators significantly influence the model ($Q_M = 7.32, df = 5, p < 0.0001$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010]. 
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 40994.51, df = 105, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

@@

#### Investigate impact of imputing Mean square errors on the meta-analysis outcomes

```{r Initiak_ReRunMetafor, eval=FALSE, include=FALSE}
# It was noticed that the meta analysis was giving different outcomes.
#     (not just small variations in estimates and p-values, but large variations)
# The following chunks investigate how much these values vary and change the outcomes
#
# This chunk creates a data.frame from the meta-analysis which was run in the previous chunk `Metafor-analysis`
# and sets the iteration number as 0
# this data.frame is appended with additional iterations to in the following chunk

summary(PM_mv_AI)

dat1 <- data.frame(
   iteration = 0,
   mods = names(coef(PM_mv_AI)),
   coefs = coef(PM_mv_AI),
   se = PM_mv_AI$se,
   zval = PM_mv_AI$zval,
   pval = PM_mv_AI$pval,
   ci.lb = PM_mv_AI$ci.lb,
   ci.ub = PM_mv_AI$ci.ub,
   row.names = NULL
)

# commented the code below so the file does not get overwritten
#write.csv(dat1, "./cache/metaIteration.csv", row.names = FALSE)
```

```{r run_iterations_asJobscript, eval=FALSE, include=FALSE}
# This chunk iterates the imputation of MSEs and reruns the metafor analysis
# each time it iterates it appends some of the output information into the "metaIteration.csv" file
# To increase the speed the calculations/iterations it is send to multiple jobs in the background.
#
# WARNING don't run a high number of iterations without increasing sys.sleep
# sys.sleep should not be reduced below 4 for 50 iterations and
# below 0.5 for 10 iterations
for (ITERATE in 1:50) {
   rstudioapi::jobRunScript(
      "./R/ReRun_metafor-grainyield.R",
      workingDir = here::here(),
      name = "metafor_iteration",
   )
   Sys.sleep(15)
}
```

```{r coef_iteration, eval=FALSE, include=FALSE}
# This chunk plots the coefficients and corresponding confidence intervals 
#     for each iteration of the metafor calculation to visualise how much variation 
#     the imputed MSE imbue on the moderator coefs

# This function alters the moderator names so they are plotted legibly in the facet strips
change_labels <- function(x1){
   x1 <- gsub(pattern = "spray_management",replacement = "",x = x1)
   x1 <- gsub(pattern = "_",replacement = "\n",x = x1)
   x1
}

metaI <- read.csv("./cache/metaIteration.csv")
metaI %>%
   ggplot(aes(x = iteration ,y = coefs,))+
   geom_ribbon(aes(ymax = ci.ub, ymin = ci.lb, x = iteration), fill = "grey70")+
   geom_line()+
   facet_grid(cols = vars(mods), labeller = labeller(mods = change_labels))+
   theme(strip.text.x = element_text(size = 9)) # change facet strip label font size

```{r pval_iteration_plot, eval=FALSE, include=FALSE}
# Plot the p - values of the spray_management moderators
# transform y axis on log scale
metaI %>%
   ggplot(aes(x = iteration , y = pval, )) +
   geom_line() +
   scale_y_log10() +
   geom_hline(
      yintercept = c(0.05, 0.001),
      colour = "grey50",
      linetype = 2
   ) +
   facet_grid(cols = vars(mods),
              labeller = labeller(mods = change_labels)) +
   theme(strip.text.x = element_text(size = 9)) +
   geom_text(
      aes(
         y = 0.05,
         x = 1,
         label = c("P = 0.05")
      ),
      size = 3,
      colour = "grey50",
      hjust = 0,
      vjust = -0.5
   )+
      geom_text(
      aes(
         y = 0.001,
         x = 1,
         label = c("P = 0.001")
      ),
      size = 3,
      colour = "grey50",
      hjust = 0,
      vjust = -0.5
   )

```{r metafor_summary}
summary(PM_mv_AI)
```

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis. The second table is in two parts(left and right). The left part indicates the level of acceptable rho (variation) when comparing treatments. All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho. `Early` and `Late_plus` treatments never occurred within the same trial which is indicated by the right side of the table.

The fixed effects, in the last table, showed that yields in single early spray treatments are not significantly different to the no spray control. 
Also commencing spray management schedules at first sign of disease (Recommended), or between 7 - 19 days after first sign (late) produced significantly higher yields compared to the no spray control. 
On average a spray schedule with two or more applications stating late (7 - 19 days after first sign of powdery mildew) produced the highest yields.

### Profile plots  
Lets inspect the profile plots to ensure the model is not over-fitted. 
We expect to see the estimate align with the peak of the curve. 
Also that the shape of the line is a curve.

@@

*****

>>>>

### Formatting Variance

```{r remove_studies}
any(is.na(slimmer_PM_dat$yield_error))

slimmer_PM_dat$vi <- slimmer_PM_dat$yield_error
```

@@

$MSE = Var + bias$  
However if our MSE was calculated from an unbiased sample we can assume $bias = 0$ and therefore:  
$MSE = Var$

To calculate the sample variance (for each treatment in the tria) from the mean square error (Trial variance) all we need to do is divide by the number of samples in each treatment.
We can show this because:  
$MSE = \frac{\sum_{y_1}^{y_n}(y_i - \overline{y})^2}{n - 1}$  
Where the sum of squares for the fungicide treatment estimator ($y$) is divided by the number of fungicide treatments ($n$).

The sample variance is a similar formula except we use $x$ instead of $y$.  
$S_{Var} = \frac{\sum_{x_1}^{x_n}(x_i - \overline{x})^2}{n - 1}$  
$x$ being each observation within the experiment and $n$ being the number of samples in the whole trial.
Therefore: 
$S_{Var} = \frac{\frac{\sum_{y_1}^{y_n}(y_i - \overline{y})^2}{n_y - 1}}{n_{x}}$  
Where $n_x$ is the number of samples within each treatment.

```{r MSE_2_SVar}
#calculation when using log of response
# slimmer_PM_dat$vi <-
#    slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
#                                   2)
#
slimmer_PM_dat$vi2 <- 
   slimmer_PM_dat$Y_Msquare / 
   slimmer_PM_dat$n

@@

We have removed the studies that don't report any variance from the analysis.

### Caluculate standardised mean differences
We may wish to undertake the meta-analysis using standardised mean differences. 
Lets calculate them and add them to the dataframe.
```{r StandardisedMeanDifference}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for(T_ref in slimmer_PM_dat$trial_ref){

@@

We can use the same sample variance because if you add or subtract from a random variable the variance does not change.

Lets simplify our data
```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
      spray_management,
      c( "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )))%>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      vi,
      vi2
   )
```

@@

@@

Lets undertake the meta-analysis, we are using the `spray_management` variable as a moderator and a interactive term to the `trial` variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.
```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management -1,
   method = "ML",
   random = ~ spray_management|trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
In this result we can see that the `Early` treatment is not significantly different to zero, which in this case is the mean of the no spray control. 
However the other treatments are significantly different to zero.
The Qm [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, show the moderators significantly influence the model ($Q_M = 85.232, df = 5, p < 0.0001$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010]. 
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 40994.51, df = 105, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

@@

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
The second table is in two parts(left and right). 
The left part, rho, is the correlation of variation between the specified treatments. 
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho. `Early` and `Late_plus` treatments never occurred within the same trial which is indicated by the right side of the table (hence the warning).

The fixed effects, in the last table, showed that yields in single early spray treatments are not significantly different to zero, or the no spray control.

### Profile plots  
Lets inspect the profile plots to ensure the model is not over-fitted. 
We expect to see the estimate align with the peak of the curve. 
Also that the shape of the line is a curve.  
Caution! this will take some time to run.

@@

*****
### Metafor comparisons
